---
title: "Data_screening"
author: "Hardy Griesbauer"
date: "06/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data screening
This document describes the process used to screen the data, fix errors, remove outliers, etc..
The data file saved at the end of this file will be in the data folder.

# Load data
```{r}
load("datRaw.RData") # note this file is saved in the data-raw folder
```

# Preliminary data screening

```{r}
# Check to make sure all species are coded correctly
unique(dat$Species)

```
Note that 'W' refers to Willow, as per discussion with Ken.  Let's remove willow from dataset.

```{r remove willow from dataset}

dat<-
  dat %>% 
  filter(Species!="W") %>%  # remove willow from dataset
  mutate(Species=fct_drop(Species)) # drop factor level

```

Other than that, all species look good.



# Screen for outliers

## DBH
```{r}
dat %>% 
  ggplot(aes(y=DBH))+
  geom_boxplot()

```

Some big trees in there!  Let's take a look at all trees over 25cm DBH

```{r}

dat[dat$DBH>25,]
```
Seem ok, although some of those trees are pretty wide for their heights.  I'll take a look at some of the plot cards to double check DBH measurements, however, this doesn't seem like a typo.

## Height
```{r}
dat %>% 
  ggplot(aes(y=Height))+
  geom_boxplot()

```

Note that there are some 'NA' values for trees that were heavy leaners (we couldn't really record a height for them)

## DBH vs Height
Let's do a scatter of DBH vs Height to make sure there are no weird DBH-Height combinations
```{r}
dat %>% 
  ggplot()+
  aes(x=Height,y=DBH)+
  geom_point()

```

Bottom portion of graph makes sense, as we generally measured trees > 4cm DBH

## Cores taken 
Make sure that trees identified as being cored make sense
```{r}
dat %>% 
  filter(Core==1) %>% # filter for trees with core identified
  group_by(Species) %>% 
  summarise(n())
```
Doesn't make sense that they would have cored At, Bl, or EP.  Let's look at those:

```{r}
dat %>% 
  filter(Core==1) %>% 
  filter(Species%in%c("Bl","At","Ep"))

```

I took a look at the Ep tree in Plot 6 of Opening 83D75-20.  They did mark a '1'.  Doesn't really make sense, but I won't change anything at this point.

## Ages missing for some blocks
```{r add ages}

# 160_D23 planted in 2002
# 83D75-20 planted in 1981
dat<-
  dat %>% 
  mutate(YearPlanted=replace(YearPlanted,Opening=="160_D23",2002)) %>% 
  mutate(YearPlanted=replace(YearPlanted,Opening=="83D75-20",1981))

```

## Change Opening number for 83D75-020
Looking at the maps, it looks like it should be 83D75-30 not 20.  Let's change that.
```{r}

dat<-
  dat %>% 
  mutate(Opening=replace(Opening,Opening=="83D75-20","83D75-30"))

```


# SAVE FILE
So far, everything looks ok.  Can put more data screening methods above as I think of them.
For now, will save data as final dataset.

```{r}
save(dat,file="../data/dat.RData")
```



