---
title: "Initial data review"
author: "Hardy Griesbauer"
date: "24/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)

# Load data
load("../../data/dat.RData")

```
## Roadmap
Ideas for analysis:
- Tree Height ~ Species + Age + Basal area + BGC Unit + Site Series + (1|Opening/Plot)
- Tree DBH ~ same variables as above
- Try the above with separate models for each species
- Generate provenance and site normals, and compare the two (generate climatic distances)

## Notes on Data
Remove Carmella block from analysis, because no western larch were planted there


## Preliminary stats

### Number of trees sampled
```{r}

dat %>% 
  group_by(Species) %>% 
  summarise(Num=n()) %>% 
  arrange(desc(Num)) %>% 
  knitr::kable()

```

## Number of Openings, Plots, and BGC units

```{r} 

dat %>% 
  group_by(BGC.unit,Opening,Plot) %>% 
  summarise(Num=n()) %>% 
  ungroup() %>% 
  group_by(BGC.unit,Opening) %>% 
  summarise(Plots=n(),
            Trees=sum(Num)) %>% 
  ungroup() %>% 
  group_by(BGC.unit) %>% 
  summarise(Openings=n(),
            Plots=sum(Plots),
            Trees=sum(Trees)) %>% 
  arrange(desc(Trees)) %>% 
  knitr::kable()
```


## Fixed effects models

Height model

```{r}
# Western larch fixed effects

height.fit<-
  dat %>% 
  filter(Species%in%c("Lw")) %>% 
  mutate(Age=2019-YearPlanted+1) %>% 
  lmer(Height~BGC.unit+SMR+(1|Opening/Plot)+(1|Age),data=.) 


# Ponderosa pine
height.fit<-
  dat %>% 
  filter(Species%in%c("Py")) %>% 
  mutate(Age=2019-YearPlanted+1) %>% 
  lmer(Height~SMR+(1|Opening/Plot),data=.) 

summary(height.fit)


```

Controlling for age and plot nested within opening, larch height is controlled by soil moisture regime, but not climatic regime.



