---
title: "Data cleaning and screening for 2019 offsite dendro data"
author: "Hardy Griesbauer"
date: "23/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(here)
library(tidyverse)
library(dplR)


# Load data
load(here("data","offsiteData.RData"))

```

# Data summaries, screening, cleaning, etc.

First, let's get a summary of number of trees in our dataset.

```{r tree data summary}



dat %>% 
  pluck("treeDat") %>% 
  group_by(Species) %>% 
  summarise(Num.Trees=n())



```

We collected `r nrow(dat$treeDat)` tree measurements in `r dat %>% pluck("treeDat") %>% distinct(Opening,Plot) %>% nrow()` plots (3.99m radius) distributed across `r dat %>% pluck("treeDat") %>% distinct(Opening) %>% nrow()` openings in the Omineca Region.

We also collected `r dat$ringDat %>% distinct(coreID) %>% nrow()` cores.

## Data screening

Let's create some boxplots of ring data to detect any outliers:

```{r boxplots}

dat %>% 
  pluck("ringDat") %>% 
  ggplot()+
  geom_boxplot(aes(y=Value,fill=varType,labels=varType))

```

We can see that earlywood and ringwidth measurements exceed 20mm in some cases.  Let's take a closer look at those:

```{r rw and ew outlier}

dat %>% 
  pluck("ringDat") %>% 
  filter(varType %in% c("ew","rw")) %>% 
  filter(Value>10) %>%  # let's look at values above 10mm
  arrange(desc(Value))

```

I checked core 5-Py-1 from Opening 3A_7.  It does indeed look like the 2002 ring was exceptionally wide.  Same with tree 411_1_A Plot 2 Tree 10.  

## Basal area increment
Unfortunately we cannot compute basal area increment, because we didn't measure the 2019 ring, which would have been partially formed at the time of sampling.  In many cases, it fell off during extraction.  This means that we cannot calculate the outermost diameter for the 2018 year ring.

## Ring patterns
Let's look at some of the annual raw ring widths:

```{r}
siteName="411_1A"
spp="Lw"
ringVar="rw"

dat %>% 
  pluck("ringDat") %>% 
  filter(Opening %in% siteName & Species %in% spp & varType %in% ringVar) %>% 
  
 
  
  # plotting
  ggplot()+
  aes(x=Year,y=Value,fill=coreID)+
  geom_line()+
  facet_wrap(~coreID)



```

It looks like there are a few trees that had multiple cores.  Let's try and screen for them>:

```{r}
dat %>% 
  pluck("ringDat") %>% 
  filter(varType=="rw") %>% 
  
  group_by(Opening,coreID) %>% 
  summarise(Num.Tree=n()) %>% 
  arrange(desc(Num.Tree))


```

# Need to come up with a function to identify cores with duplicated years
```{r}
dupList<-
  dat %>% 
  pluck("ringDat") %>% 
  filter(varType=="rw") %>% 
  group_by(Opening,coreID,Year) %>% 
  summarise(Num.Years=n()) %>% 
  filter(Num.Years>1) %>% 
  distinct(Opening,coreID)

dupList
```

### Now let's try and remove duplicates

Corkscrew block:

```{r}

opID="Corkscrew"
plotID=6


dat$treeDat %>% 
  filter(Opening==opID & Plot==plotID) %>% 
  arrange(Plot,Tree.Number) 

dat %>% 
  pluck("ringDat") %>% 
  filter(varType=="rw") %>% 
  filter(Opening==opID & Plot==plotID)



```

Looks like we have two cores for tree #3 in plot 6.  We can likely safely remove one.



------------
Block 154_75

```{r}
opID="154_75"



dat$treeDat %>% 
  filter(Opening==opID & Plot==plotID) %>% 
  arrange(Plot,Tree.Number) 

dat %>% 
  pluck("ringDat") %>% 
  filter(varType=="rw") %>% 
  filter(Opening==opID & Plot==plotID) %>% 
  group_by(Tree.Number,Year) %>% 
  summarise(n=n()) %>% 
  filter(n>1)



```

Looks like we have core 4 duplicated.  We need to remove one of them, but not sure how we would do this easily, because the data for both cores are combined.  We'll just remove both cores.

------------
Block 3A_7

```{r}
opID="3A_7"
plotID=3


dat$treeDat %>% 
  filter(Opening==opID & Plot==plotID) %>% 
  arrange(Plot,Tree.Number) 

dat %>% 
  pluck("ringDat") %>% 
  filter(varType=="rw") %>% 
  filter(Opening==opID ) %>% View()


```

We have two cores without ID - we need to remove them.


------------
Block 907_162

```{r}
opID="907_162"

dat$treeDat %>% 
  filter(Opening==opID & Plot==plotID) %>% 
  arrange(Plot,Tree.Number) 

dat %>% 
  pluck("ringDat") %>% 
  filter(varType=="rw") %>% 
  filter(Opening==opID ) %>% View()


```


We have a blank coreID in block 411_1A.  After some investigation, I think we need to remove the core.

#### Remove problematic cores

```{r}

rw<-
  dat %>% 
  pluck("ringDat") %>% 
  drop_na(coreID) %>% 
  mutate(coreID=toupper(coreID)) %>% 
  filter(!(Opening=="Corkscrew" & coreID=="6-LW-3")) %>% 
  filter(!(Opening=="154_75" & coreID=="3-LW-4"))


# Let's double check there are no empty coreID and no duplicated cores
rw %>% 
  filter(is.na(coreID))

rw %>% 
  filter(varType=="rw") %>% 
  group_by(Opening,coreID,Year) %>% 
  summarise(Num=n()) %>% 
  filter(Num>1)


```

## Extract tree data
```{r}

treeData<-
  dat %>% 
  pluck("treeDat")



```

```{r}
x<-list(rw,treeData)
save(x,file="romData.rds")


```

